PoVaTiM: A Possibilistic Valid-Time Model \\ \vspace*{0.04truein}

### 2012-04-16 ijcis-pos-model/sources/preliminaries.tex
version sent to olga and nicolas.

\subsection{Interval Evaluation by Ill-known Constraints}
The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$. In \cite{Pons2011}, the notion of an \emph{ill-known constraint} is introduced:

\begin{definition}
Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
C \triangleq (R,V)
\end{align}
Set $A$ now satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}

@@

The uncertainty about the evaluation of an ill-known value $X$ is given by:
\begin{align}
\label{eq:ill-known-value-def-by-const-unc}
X \in \Pow(U):\pi_{X}=\pi_{CC}
\end{align} 
Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demostrated in theorem \ref{th:convex-combination-ill-known-constraints}.
\end{definition}

>>>>

\subsection{\label{subsec:interval-evaluation-by-ill-known-constraints}Interval Evaluation by Ill-known Constraints}
The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$. In \cite{Pons2011}, the notion of an \emph{ill-known constraint} is introduced:

\begin{definition}
Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
\label{eq:ill-known-constraint}
C \triangleq (R,V)
\end{align}
Set $A$ now satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}

@@

The uncertainty about the evaluation of an ill-known value $X$ is given by:
\begin{align}
\label{eq:ill-known-value-def-by-const-unc}
X \in \Pow(U):\pi_{X}=\pi_{CC}
\end{align} 
Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in theorem \ref{th:convex-combination-ill-known-constraints}.
\end{definition}

### 2012-04-16 ijcis-pos-model/sources/timeRepresentation.tex
version sent to olga and nicolas.

%Example of different uses of the ill-known time value.
\begin{example}
Consider a historical database with data from diplomatic documents. The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the stimated time when the document was sent (field \emph{Date}). Table \ref{tbl:sample-time-point} contains some example data from this database. 
\end{example}

@@

In that database, for the document with ID=23454, all the dates in the domain are equally possible. Nevertheless, the document 34563 was sent in the crisp (exact) date 11/12/1204. The time for documents 12211 and 23455 are specified with several possibility distributions. The first one is also known as a fuzzy number whereas the second one is also known as a fuzzy interval.
\textcolor{red}{References to the preliminaries}.

@@

\begin{definition}
\emph{Ill-known time interval}\\
An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval that for some reason, its boudaries are not precissely known. The ill-known interval is defined by two ill-known points, namely $X$,$Y$.
In order to know if all the points in the crisp time interval $I=\left[a, b\right]$ are within the boudaries of the ill-known time interval $\left[X, Y\right]$ we define the following two ill-known constraints $C_{1}, C_{2}$ 
\begin{eqnarray}
C_1\stackrel{\triangle}{=}\left(\geq,X\right) \wedge \\
C_2\stackrel{\triangle}{=}\left(\leq,Y\right)
\end{eqnarray}
Note that the boolean operator applied to the constraints is $\bool=\wedge$.
This means that we want to know if all the points in the interval $I$ are greater or equal to $X$ and, at the same time, smaller or equal to $Y$.
\end{definition}

@@

\subsubsection{\label{subsubsec:open-interval}Open ill-known time intervals}
Quite often, the user may want to specify time intervals with open boudaries in one or both endpoints. Consider an ill-known time interval $\left[X, Y\right]$. Then it is possible to distinguish between the following two types of open intervals:
\begin{enumerate}
\item
\emph{Completely unknown time interval}: Both starting and ending points are unknown, therefore the hole interval is specified by the \emph{UNKNOWN} constant. Due to the representation, both endpoints, say $X$ and $Y$ are stored. Thus, the interval is stored as $[$ \emph{UNKNOWN}, \emph{UNKNOWN} $]$.
\item
\emph{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown. Note that due to the ill-known constraints $C_{1}, C_{2}$ it is not possible a representation like $[$ \emph{UNKNOWN} $, Y]$ or $[X, $ \emph{UNKNOWN} $]$. The solution adopted for this problem is explained in the following paragraph.
\end{enumerate}

\paragraph{Representation of semi-open time intervals}
As mentioned before, the problem resides in the representation of these kind of interval. In order to get a short representation for these intervals, two constants are defined: $FB$ \emph{From the Beggining}: it is only valid for the left value of the interval and $UC$ \emph{Until Changed} (this name is usual in the temporal database community, see ~\cite{Jensen1994}) which is only valid for the right value of the interval. Note that these two constants are aliases for a function that obtains both possibility and necessity measures.

@@

\begin{definition}
\label{def:open-func}
$Open(C)$\\
The function $\Open(C)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
The parameter $C$ is an ill-known constraint $C\stackrel{\triangle}{=}\left(R,T\right)$ as defined in equation \eqref{rand} and in~\cite{Pons2011}.
The possibility and necessity measures are defined by:
\begin{eqnarray}
\label{eq:open-pos-nec}
\Pos\left(\Open(C)\right) &=& \Big(\sup_{r\in\T \Rp\  w}\pi_{T}(w)\Big)\\
\Nec\left(\Open(C)\right) &=& \Big(\inf_{r\in\T \Rn\  w}1-\pi_{T}(w)\Big)
\end{eqnarray}
Where the values for the binary relations $\Rp$ and $\Rn$ are shown in Table \ref{tbl:open-pos-nec-rels}.
\end{definition}

>>>>

%Example of different uses of the ill-known time value.
\begin{example}
Consider a historical database with data from diplomatic documents. The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the estimated time when the document was sent (field \emph{Date}). Table \ref{tbl:sample-time-point} contains some example data from this database. 
\end{example}

@@

In that database, for the document with ID=23454, all the dates in the domain are equally possible. Nevertheless, the document 34563 was sent in the crisp (exact) date 11/12/1204. The time for documents 12211 and 23455 are specified with several possibility distributions. The first one is also known as a fuzzy number whereas the second one is also known as a fuzzy interval, as explained in Section \ref{sec:prelim}.

@@

\begin{definition}
\emph{Ill-known time interval}\\
An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval that for some reason, its boundaries are not precisely known. The ill-known interval is defined by two ill-known points, namely $X$,$Y$.
In order to know if all the points in the crisp time interval $I=\left[a, b\right]$ are within the boundaries of the ill-known time interval $\left[X, Y\right]$ we define the following two ill-known constraints $C_{1}, C_{2}$ 
\begin{eqnarray}
C_1\stackrel{\triangle}{=}\left(\geq,X\right) \wedge \\
C_2\stackrel{\triangle}{=}\left(\leq,Y\right)
\end{eqnarray}
Note that the boolean operator applied to the constraints is $\bool=\wedge$.
This means that we want to know if all the points in the interval $I$ are greater or equal to $X$ and, at the same time, smaller or equal to $Y$.
\end{definition}

@@

\subsubsection{\label{subsubsec:open-interval}Open ill-known time intervals}
Quite often, the user may want to specify time intervals with open boundaries in one or both endpoints. Consider an ill-known time interval $\left[X, Y\right]$. Then it is possible to distinguish between the following two types of open intervals:
\begin{enumerate}
\item
\emph{Completely unknown time interval}: Both starting and ending points are unknown, therefore the hole interval is specified by the \emph{UNKNOWN} constant. Due to the representation, both endpoints, say $X$ and $Y$ are stored. Thus, the interval is stored as $[$ \emph{UNKNOWN}, \emph{UNKNOWN} $]$.
\item
\emph{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown. Note that due to the ill-known constraints $C_{1}, C_{2}$ it is not possible a representation like $[$ \emph{UNKNOWN} $, Y]$ or $[X, $ \emph{UNKNOWN} $]$. The solution adopted for this problem is explained in the following paragraph.
\end{enumerate}

\paragraph{Representation of semi-open time intervals}
As mentioned before, the problem resides in the representation of these kind of interval. In order to get a short representation for these intervals, two constants are defined: $FB$ \emph{From the Beginning}: it is only valid for the left value of the interval and $UC$ \emph{Until Changed} (this name is usual in the temporal database community, see ~\cite{Jensen1994}) which is only valid for the right value of the interval. Note that these two constants are aliases for a function that obtains both possibility and necessity measures.

@@

\begin{definition}
\label{def:open-func}
$Open(C)$\\
The function $\Open(C)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
The parameter $C$ is an ill-known constraint $C\stackrel{\triangle}{=}\left(R,T\right)$ as defined in \ref{subsec:interval-evaluation-by-ill-known-constraints}, in the equation \eqref{eq:ill-known-constraint} and in~\cite{Pons2011}.
The possibility and necessity measures are defined by:
\begin{eqnarray}
\label{eq:open-pos-nec}
\Pos\left(\Open(C)\right) &=& \Big(\sup_{r\in\T \Rp\  w}\pi_{T}(w)\Big)\\
\Nec\left(\Open(C)\right) &=& \Big(\inf_{r\in\T \Rn\  w}1-\pi_{T}(w)\Big)
\end{eqnarray}
Where the values for the binary relations $\Rp$ and $\Rn$ are shown in Table \ref{tbl:open-pos-nec-rels}.
\end{definition}
